### 2013-08-27 thesis/src/chapters/chapter-1.tex
structure

\chapter{Introduction}

\section{Motivation}

\section{Main goals}

\section{Impact}

\section{Thesis structure}