### 2013-08-27 thesis/src/chapters/chapter-1.tex
Add the first version of 'main goals' & 'impact'

\chapter{Introduction}

\section{Motivation}

\section{Contributions}
The main contributions of this dissertation are as follows:
\begin{itemize}
  \item A proposal of an architecture of federated cloud computing environment, which is based on and can be viewed as a simplified version of \emph{InterCloud}
  \item The notion of considering each service model as an autonomic system
  \item The implementation of the proposed architecture using OpenNebula technology stack
\end{itemize}

\section{Impact}
We hope that the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways the following features can be implemented:
\begin{itemize}
  \item autoscaling capabilities
  \item virtual machines provisioning
  \item managing of a group of virtual machines which forms a logical entity
\end{itemize}
What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity. And finally, the idea of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
\section{Thesis structure}