### 2013-09-11 thesis/src/appendices/listings.tex
Added policy description

\chapter{Code listings}

\section{Scaling policies}
\label{app:scaling-policies}

\lstinputlisting[caption=Auto Scaling policy\, AWS EC2]{scaling-policy.xml}