### 2013-09-12 thesis/src/appendices/listings.tex
Added listing style

\chapter{Code listings}

\section{Scaling policies}
\label{app:scaling-policies}

Listings \ref{lst:scaling-ec2} illustrates XML-based policy that is used by Auto Scaling of the Amazon Web Services EC2.
\lstinputlisting[caption=Scaling policy - AWS EC2, label=lst:scaling-ec2]{appendices/scaling-policy.xml}