### 2013-12-06 thesis/src/chapters/chapter-7.tex

\chapter{Evaluation}

\chapterintro{
  This chapter contains results and discussions on the evaluation of tests run
  on the proposed solution.
}