### 2013-12-12 thesis/src/chapters/chapter-6-implementation.tex
Add requirements section

\chapter{Implementation}

\chapterintro{
  In this chapter we outline implementation details about each component of the
  proposed solution.
}

\section{Application controller}
\section{Container controller}
\section{Service controller}
\section{Cloud controller, cloud broker and client application}