### 2013-12-23 thesis/src/appendices/listings.tex
New deployment time figure

\chapter{Code listings}

\section{Service specifications}
\lstinputlisting[caption=Carina Environment Specification which was used during tests of deployment time, label=lst:carina-config]{appendices/carina-svc-config.rb}

\lstinputlisting[caption=Cloud-SAP Service Specification (without scaling policies), label=lst:service-spec-test-cost]{appendices/service-spec-test-cost.json}

\lstinputlisting[caption=Cloud-SAP Service Specification used for testing deployment time, label=lst:service-spec-test-deployment-time]{appendices/service-spec-deployment-time-test.json}

\section{Scaling policies}
\label{app:scaling-policies}

Listings \ref{lst:scaling-ec2} illustrates XML-based policy that is used by Auto Scaling of the Amazon Web Services EC2.
\lstinputlisting[caption=Scaling policy - AWS EC2, label=lst:scaling-ec2]{appendices/scaling-policy.xml}