### 2013-12-26 thesis/src/chapters/chapter-5-design.tex
Thesis: chapter 5 structure

\chapter{Design of Cloud-SAP}

\chapterintro{ This chapter introduces the high-level design of Cloud-SAP.
  Further elaboration on the layers of the proposed platform is presented
  and detailed discussion about the each layer is held.
}

\section{Requirements}
One can notice that elements that yields a solution for a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:

\begin{itemize}
	\item \emph{scalability} - ability to improve application performance by enriching resources
	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
	\item \emph{inter-cloud awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
\end{itemize}

Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarised the design choices in a context of system requirements.

\section{High-level design}
\subsection{Key elements}
In the previous chapters, various scalability models operating at different layers of abstraction were introduced: 
\begin{itemize}
	\item application platform: application platform tuning
	\item container: vertical scaling
	\item stack: horizontal scaling
	\item inter-cloud: scaling out across different cloud providers
\end{itemize}

It is vital that proposed architecture reflects that multi-layered model. Thus, we divided platform's components into a stack that is shown in figure \ref{design:csap-layers}.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{chapter-5/sap-layers}
  \end{center}
  \caption{Layered structure of Cloud-SAP}
  \label{design:csap-layers}
\end{figure}

\subsection{Autonomic components}
Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterized by an ability to adapt to a given application usage. As the previous chapter stated, this can be achieved by adding an elasticity controller to each layer. This observation is a foundation of proposed architecture.

One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}. Figure \ref{design:autonomic-component} depicts that architecture.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-5/autonomic-component}
  \end{center}
  \caption{autonomic component}
  \label{design:autonomic-component}
\end{figure}

Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}. Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling. Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralized and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.

As \cite{IBM06} states, each autonomic component has modules that are responsible for:
\begin{itemize}
	\item monitoring
	\item analysis
	\item planning
	\item action execution
\end{itemize}

While the managed component has:
\begin{itemize}
	\item sensors
	\item effectors
\end{itemize}

Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.

\subsubsection{Monitoring}

\subsubsection{Analysis}

\subsubsection{Planning}

\subsubsection{Execution}

\subsection{Cloud federation}
Opis najwyzej warstwy, ze ma brokery, wykorzystuje kilka chmur.

\subsection{Architecture overview}
Taking above-mentioned points into account, we can distinguish two pivotal components responsible for auto-scaling and cooperation in intercloud respectively. While each auto-scaling component is entirely dedicated to a given cloud provider, the intercloud module has distributed components such as client and broker and can correspond to a variety of different auto-scaling components. Figure X illustrates overall architecture.

(Diagram komponentowy)

\section{Container controller}

\section{Stack controller}

\section{Cloud controller, cloud broker and client application}

\section{Solution discussion}