### 2014-02-23 thesis/src/chapters/chapter-summary.tex
Thesis: add draft of a summary

\chapter{Summary}

\chapterintro{This chapter provides the summery of the dissertation and gives information about future work and research directions.}

\section{Conclusions}
In this dissertation we proposed an architecture of a platform that ensures satisfying Quality of Service requirements of users' applications deployed in a cloud computing environment. The key attributes of the proposed system include self-adaptability (in terms of the ability to conduct self-monitoring, analysing, planning and executing) and cloud-federation awareness. Additionally, we implemented a minimal viable product which conforms to the devised architecture and made its evaluation by performing tests which compare it to some of the currently available solutions.

The architecture proposal and its implementation clearly indicates that the goals of the dissertation stated in the first chapter have been achieved. Since the implementation is merely a proof-of-concept product, it cannot be seen as a complete or reference solution, though. Irrespective of the simplicity of the employed prediction, planning or resource mapping algorithms, promising test results were obtained. However, in order to get better results, more sophisticated algorithms and methods should be implemented in place of them.

All tests of the platform were not simulated on any software, but conducted on the real hardware forming a home-made data center. Tests were chosen and design so as to reflect at the highest possible rate real business use cases. Their results show considerable potential of the proposed solution.

TODO co z OpenVZ--KVM?

\section{Future work}
As the provided implementation cannot be regarded as a fully fledged one, there is a need to implement a system with the advanced algorithms regarding actions in a control loop of autonomic managers. This will enable to conduct more insightful tests of the architecture and prove its usefulness.

TODO za mało ..