### 2014-02-28 thesis/src/appendices/glossary.tex
Thesis: add glossary style

\newglossaryentry{perl}{name={Perl},  description={A scripting language}}  
\newglossaryentry{tex}{name={\TeX},  description={A typesetting language},sort={TeX}}  
\newglossaryentry{html}{name={html},  description={A mark up language}}